Developmental trajectories of word recognition: a priming study in bilinguals

### 2023-10-31 manuscript/manuscript.qmd
Merge

```{r tbl-vocabulary}
#| label: tbl-vocabulary
#| tbl-cap: "Participant vocabulary sizes across ages and language profiles. Vocabulary size scores are presented using two measures: L1 vocabulary size (proportion of words marked as *Understands* in the vocabulary checklist of the dominant language), and total vocabulary size (proportion of words marked as *Understands* in vocabulary checklists of both dominant and non-dominant languages). Vocabulary size scores from the recurrent participants are shown linked."
#| eval: false
vocabulary |> 
	inner_join(select(participants, child_id, session_id, age, lp),
			   by = join_by(child_id, session_id)) |> 
	pivot_longer(c(total_prop, l1_prop),
				 names_to = "measure",
				 values_to = "prop",
				 names_transform = \(x) gsub("_prop", "", x)) |> 
	mutate(lp = case_when(
		lp=="Monolingual" ~ "Monolingual (Catalan or Spanish)",
		lp=="Bilingual" ~ "Bilingual (Catalan-Spanish)",
		.default = lp) |> 
			factor(levels = c("Monolingual (English)",
							  "Monolingual (Catalan or Spanish)",
							  "Bilingual (Catalan-Spanish)")),
		measure = if_else(measure=="l1", 
						  "L1 vocabulary", 
						  "Total vocabulary"),
		is_imputed = if_else(is_imputed, "Imputed", "Observed") |> 
			factor(levels = c("Observed", "Imputed"))) |> 
	ggplot(aes(age, prop, colour = is_imputed)) +
	facet_grid(measure ~ lp) +
	geom_line(aes(group = child_id),
			  alpha = 1/2,
			  colour = "grey") +
	geom_point() +
	labs(x = "Age (months)",
		 y = "Vocabulary size (%)") +
	theme(panel.grid = element_line(linetype = "dotted",
									colour = "grey"),
		  legend.position = "right",
		  legend.title = element_blank())
```

@@

```{r fig-ptlt-recognition}
#| label: fig-ptlt-recognition-aggr
#| fig-cap: "Average probability of target looking across the time course of the trial. Points an whiskers indicate the mean and standard error of the mean."
data_time |> 
	summarise(.prop = mean(.prop),
			  .by = c(lp, timebin, session_id)) |> 
	ggplot(aes(timebin, .prop, colour = lp)) +
	facet_wrap(~lp, nrow = 1, scales = "free_x") +
	geom_line(aes(group = session_id),
			  # colour = "grey",
			  alpha = 1/10) +
	geom_hline(yintercept = 0.5, linetype = "dashed") +
	stat_summary(fun.data = "mean_se",
				 geom = "errorbar",
				 width = 0.35,
				 linewidth = 2/4) +
	stat_summary(fun = "mean",
				 geom = "point",
				 size = 1.2) +
	stat_summary(fun = "mean",
				 geom = "line",
				 linewidth = 2/4) +
	labs(x = "Time (ms)",
		 y = "PTLT") +
	scale_x_continuous(breaks = seq(0, 20, 2),
					   labels = \(x) format(x*100+300, big.mark = ",")) +
	scale_y_continuous(labels = scales::percent,
					   breaks = seq(0, 1, 0.2),
					   limits = c(0, 1)) +
	theme(panel.grid.major.y = element_line(colour = "grey",
											linetype = "dotted"),
		  legend.position = "none")
```

```{r fig-ptlt-recognition-age}
#| label: fig-ptlt-recognition-age
#| fig-cap: "Average probability of target looking across the time course of the trial. Points an whiskers indicate the mean and standard error of the mean."
data_time |> 
	summarise(.prop = mean(.prop),
			  .by = c(lp, timebin, age_group, session_id)) |> 
	ggplot(aes(timebin, .prop, colour = age_group)) +
	facet_wrap(~lp, nrow = 1, scales = "free_x") +
	# geom_rug(aes(x = timebin[is.na(.prop)])) +
	# geom_line(aes(group = session_id),
	# 		  alpha = 1/8) +
	geom_hline(yintercept = 0.5, linetype = "dashed") +
	stat_summary(fun.data = "mean_se",
				 geom = "errorbar",
				 width = 0.35,
				 linewidth = 2/4) +
	stat_summary(fun = "mean",
				 geom = "point",
				 size = 1.2) +
	stat_summary(fun = "mean",
				 geom = "line",
				 linewidth = 2/4) +
	labs(x = "Time (ms)",
		 y = "PTLT",
		 colour = "Age") +
	scale_x_continuous(breaks = seq(0, 20, 4),
					   labels = \(x) format(x*100+200, big.mark = ",")) +
	scale_y_continuous(labels = scales::percent,
					   breaks = seq(0, 1, 0.05)) +
	theme(legend.position = "top",
		  panel.grid.major.y = element_line(colour = "grey",
		  								  linetype = "dotted"))
```

```{r fig-ptlt-recognition-condition}
#| label: fig-ptlt-recognition-condition
#| fig-cap: "Average probability of target looking across the time course of the trial. Points an whiskers indicate the mean and standard error of the mean."
data_time |> 
	summarise(.prop = mean(.prop),
			  .by = c(lp, timebin, condition, session_id)) |> 
	ggplot(aes(timebin, .prop, colour = condition)) +
	facet_grid(~lp) +
	# geom_rug(aes(x = timebin[is.na(.prop)])) +
	# geom_line(aes(group = interaction(condition, session_id)),
	# 		  alpha = 1/8) +
	geom_hline(yintercept = 0.5, linetype = "dashed") +
	stat_summary(fun.data = "mean_se",
				 geom = "errorbar",
				 width = 0.35,
				 linewidth = 2/4) +
	stat_summary(fun = "mean",
				 geom = "point",
				 size = 1.2) +
	stat_summary(fun = "mean",
				 geom = "line",
				 linewidth = 2/4) +
	labs(x = "Time (ms)",
		 y = "PTLT",
		 colour = "Condition") +
	scale_x_continuous(breaks = seq(0, 20, 4),
					   labels = \(x) format(x*100+200, big.mark = ",")) +
	scale_y_continuous(labels = scales::percent,
					   breaks = seq(0, 1, 0.05)) +
	theme(legend.position = "top",
		  panel.grid.major.y = element_line(colour = "grey",
		  								  linetype = "dotted"))
```

@@

@@

epreds |> 
	ggplot(aes(timebin, .value,
			   colour = condition,
			   fill = condition)) +
	facet_wrap(~lp) +
	geom_hline(yintercept = 0,
			   linetype = "dashed") +
	geom_ribbon(aes(ymin = .lower, ymax = .upper),
				alpha = 0.5,
				linewidth = 0) +
	geom_line(linewidth = 3/4) +
	labs(x = "Time (ms)",
		 y = "Logit (Target looking)",
		 colour = "Condition",
		 fill = "Condition") +
	scale_x_continuous(breaks = seq(0, 20, 4),
					   labels = \(x) format(x*100+200, big.mark = ",")) +
	theme(legend.position = "top",
		  panel.grid.major = element_line(colour = "grey",
		  								linetype = "dotted"))
```

@@

epreds |> 
	ggplot(aes(timebin, .value,
			   colour = condition,
			   fill = condition)) +
	facet_grid(lp~voc_total) +
	geom_hline(yintercept = 0,
			   linetype = "dashed") +
	geom_ribbon(aes(ymin = .lower, ymax = .upper),
				alpha = 0.5,
				linewidth = 0) +
	geom_line(linewidth = 3/4) +
	labs(x = "Time (ms)",
		 y = "Logit (Target looking)",
		 colour = "Condition",
		 fill = "Condition") +
	scale_x_continuous(breaks = seq(0, 20, 4),
					   labels = \(x) format(x*100+200, big.mark = ",")) +
	theme(legend.position = "top",
		  panel.grid.major = element_line(colour = "grey",
		  								linetype = "dotted"))
```

>>>>

```{r tbl-vocabulary}
#| label: tbl-vocabulary
#| tbl-cap: "Participant vocabulary sizes across ages and language profiles. Vocabulary size scores are presented using two measures: L1 vocabulary size (proportion of words marked as *Understands* in the vocabulary checklist of the dominant language), and total vocabulary size (proportion of words marked as *Understands* in vocabulary checklists of both dominant and non-dominant languages). Vocabulary size scores from the recurrent participants are shown linked."
#| eval: false
vocabulary |> 
	inner_join(select(participants, child_id, session_id, age, lp),
			   by = join_by(child_id, session_id)) |> 
	pivot_longer(c(total_prop, l1_prop),
				 names_to = "measure",
				 values_to = "prop",
				 names_transform = \(x) gsub("_prop", "", x)) |> 
	mutate(lp = case_when(
		lp=="Monolingual" ~ "Monolingual (Catalan or Spanish)",
		lp=="Bilingual" ~ "Bilingual (Catalan-Spanish)",
		.default = lp) |> 
			factor(levels = c("Monolingual (English)",
							  "Monolingual (Catalan or Spanish)",
							  "Bilingual (Catalan-Spanish)")),
		measure = if_else(measure=="l1", 
						  "L1 vocabulary", 
						  "Total vocabulary"),
		is_imputed = if_else(is_imputed, "Imputed", "Observed") |> 
			factor(levels = c("Observed", "Imputed"))) |> 
	ggplot(aes(age, prop, colour = is_imputed)) +
	facet_grid(measure ~ lp) +
	geom_line(aes(group = child_id),
			  alpha = 1/2,
			  colour = "grey") +
	geom_point() +
	labs(x = "Age (months)",
		 y = "Vocabulary size (%)") +
	theme(panel.grid = element_line(linetype = "dotted",
									colour = "grey"),
		  legend.position = "right",
		  legend.title = element_blank())

ggsave(file.path("img", "vocabulary.jpeg"),
	   dpi = 800, height = 8, height = 4)
```

@@

```{r fig-ptlt-recognition}
#| label: fig-ptlt-recognition-aggr
#| fig-cap: "Average probability of target looking across the time course of the trial. Points an whiskers indicate the mean and standard error of the mean."
data_time |> 
	summarise(.prop = mean(.prop),
			  .by = c(lp, timebin, session_id)) |> 
	ggplot(aes(timebin, .prop, colour = lp)) +
	facet_wrap(~lp, nrow = 1, scales = "free_x") +
	geom_line(aes(group = session_id),
			  # colour = "grey",
			  alpha = 1/10) +
	geom_hline(yintercept = 0.5, linetype = "dashed") +
	stat_summary(fun.data = "mean_se",
				 geom = "errorbar",
				 width = 0.35,
				 linewidth = 2/4) +
	stat_summary(fun = "mean",
				 geom = "point",
				 size = 1.2) +
	stat_summary(fun = "mean",
				 geom = "line",
				 linewidth = 2/4) +
	labs(x = "Time (ms)",
		 y = "PTLT") +
	scale_x_continuous(breaks = seq(0, 20, 2),
					   labels = \(x) format(x*100+200, big.mark = ",")) +
	scale_y_continuous(labels = scales::percent,
					   breaks = seq(0, 1, 0.2),
					   limits = c(0, 1)) +
	theme(panel.grid.major.y = element_line(colour = "grey",
											linetype = "dotted"),
		  legend.position = "none")

ggsave(file.path("img", "recognition.jpeg"),
	   height = 4, width = 11,
	   dpi = 800)
```

```{r fig-ptlt-recognition-age}
#| label: fig-ptlt-recognition-age
#| fig-cap: "Average probability of target looking across the time course of the trial. Points an whiskers indicate the mean and standard error of the mean."
data_time |> 
	summarise(.prop = mean(.prop),
			  .by = c(lp, timebin, age_group, session_id)) |> 
	ggplot(aes(timebin, .prop, colour = age_group)) +
	facet_wrap(~lp, nrow = 1, scales = "free_x") +
	# geom_rug(aes(x = timebin[is.na(.prop)])) +
	# geom_line(aes(group = session_id),
	# 		  alpha = 1/8) +
	geom_hline(yintercept = 0.5, linetype = "dashed") +
	stat_summary(fun.data = "mean_se",
				 geom = "errorbar",
				 width = 0.35,
				 linewidth = 2/4) +
	stat_summary(fun = "mean",
				 geom = "point",
				 size = 1.2) +
	stat_summary(fun = "mean",
				 geom = "line",
				 linewidth = 2/4) +
	labs(x = "Time (ms)",
		 y = "PTLT",
		 colour = "Age") +
	scale_x_continuous(breaks = seq(0, 20, 4),
					   labels = \(x) format(x*100+200, big.mark = ",")) +
	scale_y_continuous(labels = scales::percent,
					   breaks = seq(0, 1, 0.05)) +
	theme(legend.position = "top",
		  panel.grid.major.y = element_line(colour = "grey",
		  								  linetype = "dotted"))

ggsave(file.path("img", "recognition-age.jpeg"),
	   height = 4, width = 11,
	   dpi = 800)
```

```{r fig-ptlt-recognition-condition}
#| label: fig-ptlt-recognition-condition
#| fig-cap: "Average probability of target looking across the time course of the trial. Points an whiskers indicate the mean and standard error of the mean."
data_time |> 
	summarise(.prop = mean(.prop),
			  .by = c(lp, timebin, condition, session_id)) |> 
	ggplot(aes(timebin, .prop, colour = condition)) +
	facet_grid(~lp) +
	# geom_rug(aes(x = timebin[is.na(.prop)])) +
	# geom_line(aes(group = interaction(condition, session_id)),
	# 		  alpha = 1/8) +
	geom_hline(yintercept = 0.5, linetype = "dashed") +
	stat_summary(fun.data = "mean_se",
				 geom = "errorbar",
				 width = 0.35,
				 linewidth = 2/4) +
	stat_summary(fun = "mean",
				 geom = "point",
				 size = 1.2) +
	stat_summary(fun = "mean",
				 geom = "line",
				 linewidth = 2/4) +
	labs(x = "Time (ms)",
		 y = "PTLT",
		 colour = "Condition") +
	scale_x_continuous(breaks = seq(0, 20, 4),
					   labels = \(x) format(x*100+200, big.mark = ",")) +
	scale_y_continuous(labels = scales::percent,
					   breaks = seq(0, 1, 0.05)) +
	theme(legend.position = "top",
		  panel.grid.major.y = element_line(colour = "grey",
		  								  linetype = "dotted"))

ggsave(file.path("img", "recognition-condition.jpeg"),
	   height = 4, width = 11,
	   dpi = 800)
```

@@

ggsave(file.path("img", "model-agg.jpeg"),
	   height = 7.5, width = 10,
	   dpi = 800)
```

@@

epreds |> 
	ggplot(aes(timebin, .value,
			   colour = condition,
			   fill = condition)) +
	facet_wrap(~lp) +
	geom_hline(yintercept = 0,
			   linetype = "dashed") +
	geom_ribbon(aes(ymin = .lower, ymax = .upper),
				alpha = 0.5,
				linewidth = 0) +
	geom_line(linewidth = 3/4) +
	labs(x = "Time (ms)",
		 y = "Logit (Target looking)",
		 colour = "Condition",
		 fill = "Condition") +
	scale_x_continuous(breaks = seq(0, 20, 4),
					   labels = \(x) format(x*100+200, big.mark = ",")) +
	theme(legend.position = "top",
		  panel.grid.major = element_line(colour = "grey",
		  								linetype = "dotted"))

ggsave(file.path("img", "model-time.jpeg"),
	   height = 4, width = 10,
	   dpi = 800)
```

@@

epreds |> 
	ggplot(aes(timebin, .value,
			   colour = condition,
			   fill = condition)) +
	facet_grid(lp~voc_total) +
	geom_hline(yintercept = 0,
			   linetype = "dashed") +
	geom_ribbon(aes(ymin = .lower, ymax = .upper),
				alpha = 0.5,
				linewidth = 0) +
	geom_line(linewidth = 3/4) +
	labs(x = "Time (ms)",
		 y = "Logit (Target looking)",
		 colour = "Condition",
		 fill = "Condition") +
	scale_x_continuous(breaks = seq(0, 20, 4),
					   labels = \(x) format(x*100+200, big.mark = ",")) +
	theme(legend.position = "top",
		  panel.grid.major = element_line(colour = "grey",
		  								linetype = "dotted"))

ggsave(file.path("img", "model-time-voc-total.jpeg"),
	   height = 5.5, width = 10,
	   dpi = 800)
```
